---
title: "Bdepletion"
author: "Yuezhe Li"
date: "2/16/2022"
output: html_document
---

```{r}
rm(list = ls())

library(tidyverse)
library(mrgsolve)
library(gridExtra)
library(grid)

setwd(dirname(rstudioapi::getSourceEditorContext()$path)) # set the working directory at the folder that contains the script

source('DepletionParam.r')

basemus <- mread("Bcell_mus2") # this is the base model for mouse

basehomo <- mread("Bcell_mus4") # this is the base model for human
```
# Simulation for mouse

## depleting both splenic and circulating B cells

```{r}
ssmus <- basemus %>% mrgsim(end = 150, delta = 1) %>% as_tibble() %>% tail(n = 1) %>% mutate(time = -1)

print(ssmus$BMrec/2000)  # B cell count per uL blood

ssbasemus <- basemus %>% init(ssmus)

## 100% effective depletion
sim11 <- ssbasemus %>% init(Bt = 0, BMspl = 0, BMrec = 0) %>% mrgsim(end = 140) %>% as.tibble()

## 80% effective depletion
sim12 <- ssbasemus %>% init(Bt = ssmus$Bt * 0.2, BMspl = ssmus$BMspl * 0.2, BMrec = ssmus$BMrec * 0.2) %>% mrgsim(end = 140) %>% as.tibble()


plot1 = ggplot() + 
  geom_line(data = rbind(ssmus, sim11), aes(x = time/7, y = BMrec/2000, color = '100%')) + 
  geom_line(data =  rbind(ssmus, sim12), aes(x = time/7, y = BMrec/2000, color = '80%')) + 
  labs(y = 'B cells per uL blood', x = 'time (weeks)', color = 'depletion percentage') + theme_bw() + 
  ggtitle('mouse, depleting splenic + circulating B cells')

```


## depleting circulating B cells

```{r}
ssmus <- basemus %>% mrgsim(end = 150, delta = 1) %>% as_tibble() %>% tail(n = 1) %>% mutate(time = -1)

ssbasemus <- basemus %>% init(ssmus)

## 100% effective depletion
sim21 <- ssbasemus %>% init(BMrec = 0) %>% mrgsim(end = 14) %>% as.tibble()

## 80% effective depletion
sim22 <- ssbasemus %>% init(BMrec = ssmus$BMrec * 0.2) %>% mrgsim(end = 14) %>% as.tibble()

plot2 = ggplot() + 
  geom_line(data = rbind(ssmus, sim21), aes(x = time/7, y = BMrec/2000, color = '100%')) + 
  geom_line(data =  rbind(ssmus, sim22), aes(x = time/7, y = BMrec/2000, color = '80%')) + 
  labs(y = 'B cells per uL blood', x = 'time (weeks)', color = 'depletion percentage') + theme_bw() + 
  ggtitle('mouse, depleting only circulating B cells')

```

## depleting splenic + circulating + immature B cells, and half propreB cells

```{r}
ssmus <- basemus %>% mrgsim(end = 150, delta = 1) %>% as_tibble() %>% tail(n = 1) %>% mutate(time = -1)

## 100% effective depletion
sim31 <- basemus %>% init(Bt = 0, BMspl = 0, BMrec = 0, Bi = 0, Boe = ssmus$Boe * 0.5) %>% mrgsim(end = 140) %>% as.tibble()

## 80% effective depletion
sim32 <- basemus %>% init(Bt = ssmus$Bt * 0.2, BMspl = ssmus$BMspl * 0.2, BMrec = ssmus$BMrec * 0.2, Bi = ssmus$Bi * 0.2, Boe = ssmus$Boe * 0.5) %>% mrgsim(end = 140) %>% as.tibble()

plot3 = ggplot() + 
  geom_line(data = rbind(ssmus, sim31), aes(x = time/7, y = BMrec/2000, color = '100%')) + 
  geom_line(data =  rbind(ssmus, sim32), aes(x = time/7, y = BMrec/2000, color = '80%')) + 
  labs(y = 'B cells per uL blood', x = 'time (weeks)', color = 'depletion percentage') + theme_bw() + 
  ggtitle('mouse, depleting splenic + circulating + immature & half of propre B cells')

```

## depleting splenic + circulating + immature B cells

```{r}
ssmus <- basemus %>% mrgsim(end = 150, delta = 1) %>% as_tibble() %>% tail(n = 1) %>% mutate(time = -1)

## 100% effective depletion
sim41 <- basemus %>% init(Bt = 0, BMspl = 0, BMrec = 0, Bi = 0, Boe = ssmus$Boe * 0.5) %>% mrgsim(end = 140) %>% as.tibble()

## 80% effective depletion
sim42 <- basemus %>% init(Bt = ssmus$Bt * 0.2, BMspl = ssmus$BMspl * 0.2, BMrec = ssmus$BMrec * 0.2, Bi = ssmus$Bi * 0.2, Boe = ssmus$Boe * 0.5) %>% mrgsim(end = 140) %>% as.tibble()

plot4 = ggplot() + 
  geom_line(data = rbind(ssmus, sim41), aes(x = time/7, y = BMrec/2000, color = '100%')) + 
  geom_line(data =  rbind(ssmus, sim42), aes(x = time/7, y = BMrec/2000, color = '80%')) + 
  labs(y = 'B cells per uL blood', x = 'time (weeks)', color = 'depletion percentage') + theme_bw() + 
  ggtitle('mouse, depleting splenic + circulating + immature B cells')

```

# Simulation for human

## depleting both splenic and circulating B cells

```{r}
sshomo <- basehomo %>% mrgsim(end = 150, delta = 1) %>% as_tibble() %>% tail(n = 1) %>% mutate(time = -1)

ssbasehomo <- basehomo %>% init(sshomo)

## 100% effective depletion
sim101 <- ssbasehomo %>% init(Bt = 0, BMspl = 0, BMrec = 0) %>% mrgsim(end = 140) %>% as.tibble()

## 80% effective depletion
sim102 <- ssbasehomo %>% init(Bt = sshomo$Bt * 0.2, BMspl = sshomo$BMspl * 0.2, BMrec = sshomo$BMrec * 0.2) %>% mrgsim(end = 140) %>% as.tibble()


plot10 = ggplot() + 
  geom_line(data = rbind(sshomo, sim101), aes(x = time/7, y = BMrec/5e6, color = '100%')) + 
  geom_line(data =  rbind(sshomo, sim102), aes(x = time/7, y = BMrec/5e6, color = '80%')) + 
  labs(y = 'B cells per uL blood', x = 'time (weeks)', color = 'depletion percentage') + theme_bw() + 
  ggtitle('human, depleting splenic + circulating B cells')

```
## depleting circulating B cells

```{r}
sshomo <- basehomo %>% mrgsim(end = 150, delta = 1) %>% as_tibble() %>% tail(n = 1) %>% mutate(time = -1)

ssbasehomo <- basehomo %>% init(sshomo)

## 100% effective depletion
sim201 <- ssbasehomo %>% init(BMrec = 0) %>% mrgsim(end = 14) %>% as.tibble()

## 80% effective depletion
sim202 <- ssbasehomo %>% init(BMrec = sshomo$BMrec * 0.2) %>% mrgsim(end = 14) %>% as.tibble()

plot20 = ggplot() + 
  geom_line(data = rbind(sshomo, sim201), aes(x = time/7, y = BMrec/5e6, color = '100%')) + 
  geom_line(data =  rbind(sshomo, sim202), aes(x = time/7, y = BMrec/5e6, color = '80%')) + 
  labs(y = 'B cells per uL blood', x = 'time (weeks)', color = 'depletion percentage') + theme_bw() + 
  ggtitle('human, depleting only circulating B cells')

```

## depleting splenic + circulating + immature B cells, and half propreB cells

```{r}
sshomo <- basehomo %>% mrgsim(end = 150, delta = 1) %>% as_tibble() %>% tail(n = 1) %>% mutate(time = -1)

## 100% effective depletion
sim301 <- basehomo %>% init(Bt = 0, BMspl = 0, BMrec = 0, Bi = 0, Boe = sshomo$Boe * 0.5) %>% mrgsim(end = 140) %>% as.tibble()

## 80% effective depletion
sim302 <- basehomo %>% init(Bt = sshomo$Bt * 0.2, BMspl = sshomo$BMspl * 0.2, BMrec = sshomo$BMrec * 0.2, Bi = sshomo$Bi * 0.2, Boe = sshomo$Boe * 0.5) %>% mrgsim(end = 140) %>% as.tibble()

plot30 = ggplot() + 
  geom_line(data = rbind(sshomo, sim301), aes(x = time/7, y = BMrec/5e6, color = '100%')) + 
  geom_line(data =  rbind(sshomo, sim302), aes(x = time/7, y = BMrec/5e6, color = '80%')) + 
  labs(y = 'B cells per uL blood', x = 'time (weeks)', color = 'depletion percentage') + theme_bw() + 
  ggtitle('mouse, depleting splenic + circulating + immature & half of propre B cells')
```

## depleting splenic + circulating + immature B cells

```{r}
sshomo <- basehomo %>% mrgsim(end = 150, delta = 1) %>% as_tibble() %>% tail(n = 1) %>% mutate(time = -1)

## 100% effective depletion
sim401 <- basehomo %>% init(Bt = 0, BMspl = 0, BMrec = 0, Bi = 0, Boe = sshomo$Boe) %>% mrgsim(end = 140) %>% as.tibble()

## 80% effective depletion
sim402 <- basehomo %>% init(Bt = sshomo$Bt * 0.2, BMspl = sshomo$BMspl * 0.2, BMrec = sshomo$BMrec * 0.2, Bi = sshomo$Bi * 0.2, Boe = sshomo$Boe) %>% mrgsim(end = 140) %>% as.tibble()

plot40 = ggplot() + 
  geom_line(data = rbind(sshomo, sim401), aes(x = time/7, y = BMrec/5e6, color = '100%')) + 
  geom_line(data =  rbind(sshomo, sim402), aes(x = time/7, y = BMrec/5e6, color = '80%')) + 
  labs(y = 'B cells per uL blood', x = 'time (weeks)', color = 'depletion percentage') + theme_bw() + 
  ggtitle('mouse, depleting splenic + circulating + immature B cells')
```
# Check all the plots

```{r}
library(mrggsave)

mrggsave(list(plot1, plot2, plot3, plot4, plot10, plot20, plot30, plot40), dir = "../Bcell", width = 7, height = 3, script = 'Bdepletion.Rmd')
```
